Trump vs. Harris: A Data-Driven Forecast modeling the 2024 US Presidential Elections

### 2024-11-04 paper/paper.qmd
Added citations to appendix

\[WILL CHANGE THE FORMAT OF THESE GRAPHS AND CAPTION/DESCRIPTION WILL BE ADDED\]

@@

```{r}
#| label: fig-samplesize
#| fig-cap: [Need Edit]
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```{r}
#| label: fig-percentage
#| fig-cap: [Need Edit]
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(cleaned_data, aes(x = Percentage)) +
  geom_histogram(bins = 30,
                 fill = "orange",
                 color = "black",
                 alpha = 0.7) +
  labs(title = "Distribution of Percentage", 
       x = "Percentage",
       y = "Frequency") +
  theme_minimal()
```

```{r}
#| label: fig-candidatepercentage
#| fig-cap: [Need Edit]
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

will update these box plots to different visuals due to box plots not showing the full dataset! (#Note to team: updated to violin plots to show distribution and selected only trump and harris and another one for all the candidates to show distribution of dataset)

```{r}
#| label: fig-statepercentage
#| fig-cap: [Need Edit]
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(cleaned_data, aes(x = State, y = Percentage)) +
  geom_boxplot() +
  labs(title = "Percentage by State", x = "State", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-corrmatrix
#| fig-cap: [Need Edit]
#| echo: false
#| eval: true
#| warning: false
#| message: false
numerical_data <- cleaned_data %>%
  select(PollScore, SampleSize, Percentage)

@@

The New York Times/Siena College Poll was ranked the most accurate political pollster in the US following the 2022 midterm elections. They received the max 538 rating of 3 stars, one of only four pollsters to do so. They also had a pollscore of -1.5, meaning they had the least amount of error and bias attributed to them of the 282 pollsters ranked by 538.

The Times/Siena Poll targets registered voters in the United States, emphasizing those in battleground states during presidential elections. Since the US decides it's President based on the Electoral College rather than the national popular vote, this is an important focus in order to correctly determine election outcomes. The sampling frame is derived from a voter file which containing lists of 200 million registered voters as well as their demographic information and prior voting behavior data.

Each poll typically surveys around 1,000 respondents, which is deemed sufficient for achieving a margin of sampling error of approximately 3-4 percentage points. The Times found that tripling the sample size would only reduce the margin of error by a percent or two. This sample size allows for a balance between representativeness and efficiency.

Respondents are recruited through live phone interviews conducted by trained interviewers from call centers in various states. The polling effort includes both landline and cellphone calls, with over 90% of respondents being reached via cellphones.

Times/Siena uses a stratified random sampling technique, where the sample is divided into strata based on demographics such as age, race, gender, and party affiliation. This helps to ensure proportional representation from each group. Stratification helps in achieving a sample that mirrors the voting population, minimizing biases related to demographic variables (i.e. uneducated voters are less likely to respond to polls). However, it faces limitations as stratification increases the complexity of the sampling process, and if strata are not adequately represented in the sample, the results may still skew. Times/Siena combats this by applying weighting adjustments post-survey to correct for any demographic discrepancies.

Given the declining response rates typical of phone surveys, the Times/Siena Poll faces challenges in reaching respondents. Approximately 2% of those contacted ultimately participate. This pollster uses follow-up calls to make multiple attempts to reach individuals who initially do not respond. In 2022, they conducted an experiment to find out if responses were skewed by non-response bias. This was done via a mail survey that offered a reward of up to \$25 for a response. This saw a 30% response rate, which is significantly larger than the 2% response rate Times/Siena gets over phone. Although there were differences between the types of respondents you get between the two methodologies, there was not a meaningful difference in who these respondants claim they would vote for.

The questionnaires used by Times/Siena are designed to capture a broad range of opinions while remaining concise. Interviews are kept under 15 minutes to reduce survey fatigue and maximize response rates. This pollster prioritizes making sure every viewpoint is accurately represented in order to make the questions fair and ensure that respondents are not being swayed in a particular direction. This means the questions must remain objective and clear. However, a survey with limited response options and questions may mean a respondent's view is not captured in its full breadth. It also means that more nuanced opinions may be overlooked in favor of broader trends.

@@

To maintain a national margin of error of ±4%, the total sample size will be 1,000 respondents per state. Swing states will be oversampled, with over 2,000 respondents to achieve a margin of error of ±2%. Given an estimated response rate of 2%, we aim to recruit approximately 50,000 potential respondents per state, with a target of 100,000 in battleground states.

@@

1.  Survey Development (\$7,000):
    -   The \$7,000 covers the costs associated with designing the survey, including the creation of questions, structuring the survey for clarity and flow, and ensuring that it meets methodological standards. It may also include hiring experts in survey design or consulting with statisticians to validate the survey framework and ensure questions are unbiased. Additionally, while Google Forms is free, this budget may account for platform costs if needed.
2.  Accurate Sampling (\$10,000):
    -   The sampling budget is allocated for the costs associated with selecting a representative sample of voters. This includes expenses related to stratified random sampling, where the sample is segmented by various demographics and state representation. In order to ensure each demographic share is accurately represented, we may need to purchase access to demographic data or employ statisticians to ensure that the sampling aligns with the population of each state.
3.  Recruiting Respondents (\$45,000):
    -   The majority of the budget is dedicated to recruiting respondents, which is necessary for achieving a large and demographically representative sample. This cost will cover payments to online panel providers which charge fees for access to their respondent pools. Additionally, this budget will cover costs related to conducting state-specific phone surveys and IVR calls, such as hiring call center staff or IVR technology.
4.  Respondent Incentives (\$28,000):
    -   To encourage participation, we plan to run state-level sweepstakes, where respondents have a chance to win prizes for completing the survey. A \$500 prize will be awarded to one winner in each state (+ Washington DC). The remaining budget will cover administrative expenses associated with running the sweepstakes.
5.  Data Analysis and Reporting (\$10,000):
    -   This portion of the budget is reserved for the costs related to analyzing the collected data and reporting the findings. This may include hiring data analysts or consultants who specialize in statistical analysis and survey data interpretation. The funds will also cover software tools necessary for data analysis, as well as the creation of visualizations and reports to communicate the results effectively.

>>>>

@@

```{r}
#| label: fig-samplesize
#| fig-cap: The distribution of sample sizes used in each poll
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```{r}
#| label: fig-percentage
#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(cleaned_data, aes(x = Percentage)) +
  geom_histogram(bins = 30,
                 fill = "orange",
                 color = "black",
                 alpha = 0.7) +
  labs(title = "Distribution of Percentage", 
       x = "Percentage",
       y = "Frequency") +
  theme_minimal()
```

```{r}
#| label: fig-candidatepercentage
#| fig-cap: violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

```{r}
#| label: fig-statepercentage
#| fig-cap: The percentage of support seen by state
#| echo: false
#| eval: true
#| warning: false
#| message: false
ggplot(cleaned_data, aes(x = State, y = Percentage)) +
  geom_boxplot() +
  labs(title = "Percentage by State", x = "State", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6)) 
```

```{r}
#| label: fig-corrmatrix
#| fig-cap: correlation between PollScore, the sample size of the poll, and the percentage of support for a candidate
#| echo: false
#| eval: true
#| warning: false
#| message: false
numerical_data <- cleaned_data %>%
  select(PollScore, SampleSize, Percentage)

@@

The New York Times/Siena College Poll was ranked the most accurate political pollster in the US following the 2022 midterm elections [@PollsterRating]. They received the max 538 rating of 3 stars, one of only four pollsters to do so. They also had a pollscore of -1.5, meaning they had the least amount of error and bias attributed to them of the 282 pollsters ranked by 538 [@PollsterRating].

The Times/Siena Poll targets registered voters in the United States, emphasizing those in battleground states during presidential elections. Since the US decides it's President based on the Electoral College rather than the national popular vote, this is an important focus in order to correctly determine election outcomes. The sampling frame is derived from a voter file which containing lists of 200 million registered voters as well as their demographic information and prior voting behavior data [@TimesSienaPoll].

Each poll typically surveys around 1,000 respondents, which is deemed sufficient for achieving a margin of sampling error of approximately 3-4 percentage points. The Times found that tripling the sample size would only reduce the margin of error by a percent or two. This sample size allows for a balance between representation and efficiency [@TimesSienaPoll].

Respondents are recruited through live phone interviews conducted by trained interviewers from call centers in various states. The polling effort includes both landline and cellphone calls, with over 90% of respondents being reached via cellphones [@TimesSienaPoll].

Times/Siena uses a stratified random sampling technique, where the sample is divided into strata based on demographics such as age, race, gender, and party affiliation. This helps to ensure proportional representation from each group. Stratification helps in achieving a sample that mirrors the voting population, minimizing biases related to demographic variables (i.e. uneducated voters are less likely to respond to polls). However, it faces limitations as stratification increases the complexity of the sampling process, and if strata are not adequately represented in the sample, the results may still skew. Times/Siena combats this by applying weighting adjustments post-survey to correct for any demographic discrepancies [@TimesSienaPoll].

Given the declining response rates typical of phone surveys, the Times/Siena Poll faces challenges in reaching respondents. Approximately 2% of those contacted ultimately participate. This pollster uses follow-up calls to make multiple attempts to reach individuals who initially do not respond [@TimesSienaPoll]. In 2022, they conducted an experiment to find out if responses were skewed by non-response bias [@NonResponseBias]. This was done via a mail survey that offered a reward of up to \$25 for a response. This saw a 30% response rate, which is significantly larger than the 2% response rate Times/Siena gets over phone. Although there were differences between the types of respondents you get between the two methodologies, there was not a meaningful difference in who these respondants claim they would vote for [@NonResponseBias].

The questionnaires used by Times/Siena are designed to capture a broad range of opinions while remaining concise. Interviews are kept under 15 minutes to reduce survey fatigue and maximize response rates [@TimesSienaPoll]. This pollster prioritizes making sure every viewpoint is accurately represented in order to make the questions fair and ensure that respondents are not being swayed in a particular direction. This means the questions must remain objective and clear. However, a survey with limited response options and questions may mean a respondent's view is not captured in its full breadth. It also means that more nuanced opinions may be overlooked in favor of broader trends.

@@

To maintain a national margin of error of ±4%, based on the experiment conducted by Times/Siena, the total sample size will be 1,000 respondents per state [@TimesSienaPoll]. Swing states will be oversampled, with over 2,000 respondents to achieve a margin of error of ±2%. Given an estimated response rate of 2%, we aim to recruit approximately 50,000 potential respondents per state, with a target of 100,000 in battleground states.

@@

1.  Survey Development: \$7,000
    -   The \$7,000 covers the costs associated with designing the survey, including the creation of questions, structuring the survey for clarity and flow, and ensuring that it meets methodological standards. It may also include hiring experts in survey design or consulting with statisticians to validate the survey framework and ensure questions are unbiased. Additionally, while Google Forms is free, this budget may account for platform costs if needed.
2.  Accurate Sampling: \$10,000
    -   The sampling budget is allocated for the costs associated with selecting a representative sample of voters. This includes expenses related to stratified random sampling, where the sample is segmented by various demographics and state representation. In order to ensure each demographic share is accurately represented, we may need to purchase access to demographic data or employ statisticians to ensure that the sampling aligns with the population of each state.
3.  Recruiting Respondents: \$45,000
    -   The majority of the budget is dedicated to recruiting respondents, which is necessary for achieving a large and demographically representative sample. This cost will cover payments to online panel providers which charge fees for access to their respondent pools. Additionally, this budget will cover costs related to conducting state-specific phone surveys and IVR calls, such as hiring call center staff or IVR technology.
4.  Respondent Incentives: \$28,000
    -   To encourage participation, we plan to run state-level sweepstakes, where respondents have a chance to win prizes for completing the survey. A \$500 prize will be awarded to one winner in each state (+ Washington DC). The remaining budget will cover administrative expenses associated with running the sweepstakes.
5.  Data Analysis and Reporting: \$10,000
    -   This portion of the budget is reserved for the costs related to analyzing the collected data and reporting the findings. This may include hiring data analysts or consultants who specialize in statistical analysis and survey data interpretation. The funds will also cover software tools necessary for data analysis, as well as the creation of visualizations and reports to communicate the results effectively.
